---
title: "My R Markdown Primer"
author: "Derek"
date: "2026-04-13"
output: html_document
---

## What is R Markdown

Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. R Markdown documents are fully reproducible. 

Here is a formula for a simple linear model:

$$y_i = \beta_0 + \beta_1 x_i + \epsilon_i$$

In this section, we will load our libraries (ggplot). We will then create a dataset for all ages in a family and plot that data.

```{r load-data, message = FALSE, warning = FALSE}
library(ggplot2)

mother <- 32
father <- 31
son <- 5
daughter <- 3
uncle <- 28
aunt <- 24
grandma <- 64
grandpa <- 70

familydf <- data.frame(FamilyMembers = c("mother", "father", "son", "daughter", "uncle", "aunt", "grandma", "grandpa"), 
           Age = c(mother, father, son, daughter, uncle, aunt, grandma, grandpa))

```


```{r load-plot, message = FALSE, warning = FALSE}

familyplt <- ggplot(familydf, aes(Age, FamilyMembers)) + 
  theme_classic() +
  geom_segment(aes(x = 0, xend = Age, y = reorder(FamilyMembers, Age), yend = reorder(FamilyMembers, Age))) +
  labs(title = "Family Age Distribution", 
       subtitle = "Ranked from youngest to oldest", x = "Age (Years)", y = NULL
       ) +
  geom_point()

familyplt


```


Writing in R Markdown is very different from a regular report. R Markdown allows for dynamic reports that have visualizations and coding embedded into the report itself. It can be used to explain or demonstrate topics much more efficiently than a normal report is capable of. Everything in R Markdown can be reproduced very easily. 
